---
title: "Data Visulization Project on Movies Shot in Toronto"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

```{r,warning=FALSE, echo=FALSE, include=FALSE}
#import libraries 
library(tidyverse)
library(readr)
suppressWarnings(library(dplyr))
library(ggplot2)
library(janitor)
library(stringr)
library(quantmod)
```



```{r, echo = FALSE}
# Glimpse Data 
tor_movies <- read.csv("toronto-movies.csv")
tor_movies_clean <- tor_movies %>% janitor::clean_names()
```

 
```{r, echo= FALSE}
# Clean NAs and change to proper class
tor_movies_subset <- 
  tor_movies_clean %>% 
  select(awards, genre, imdb_rating, metascore, title, year) %>%
  mutate_at(vars(awards, genre, imdb_rating, metascore, title), na_if, "N/A") %>% 
  mutate_at(vars(imdb_rating, metascore,year), as.numeric) %>% 
  mutate_if(is.numeric, replace_na, 0) 

```



```{r, echo= FALSE}
# Plot the total number of movies by year
summary <- 
  tor_movies_subset %>% 
  group_by(genre) %>% 
  summarise(count_num = n(), .groups = 'drop') 

#Notice that the dataset isn't tidy lots of movies with multiple genres

summary_2 <- tor_movies_subset %>% 
  group_by(year) %>% 
  summarise(count = n(), .groups = "drop")
  


```

```{r, echo= FALSE}
# Importing Data from https://ca.investing.com/currencies/cad-usd-historical-data

library(ggthemes)
# import data (1 CAD = ?USD)
x_rate <- read.csv("CAD_USD Historical Data.csv")
colnames(x_rate) <- c("date", "rate", "open", "high", "low", "change")

# extracting only the rate and date 
x_rate_subset <- x_rate %>% 
  select(date,rate) 

# Noticed that date is in class character, need to change it to either date or numeric 

x_rate_subset$date <- as.Date.character(x_rate_subset$date, format = "%b %d, %Y")


# Now, after changing the date into class "Date", try to group them into groups by year so we can match the exchange rate with the toronto movie data
x_rate_subset <- x_rate_subset %>% 
  mutate(year = format(date, "%Y")) %>% 
  group_by(year) %>% 
  summarise(avg_rate = mean(rate), .groups = "drop") 

x_rate_subset$year <- as.numeric(x_rate_subset$year)



```


```{r, warning= FALSE, echo= FALSE, message= FALSE}
# merge exchange rate with toronto movies 
full_data <- x_rate_subset %>% 
  full_join(summary_2)


### Data Visualization of the full data 

# full_data %>%
#   ggplot(aes(x = year)) +
#   geom_bar(aes(y = count), stat = "Identity") +
#   geom_line(aes(y = 15*avg_rate)) +
#   scale_y_continuous(sec.axis = sec_axis(~./15))


# Data_visualization without the movie from 1933
Without_outlier <- full_data %>% 
  filter(year!= 1933)

### plot it out 

# Without_outlier %>% 
#   ggplot(aes(x = year)) + 
#   geom_bar(aes(y = count), stat = "Identity") + 
#   geom_line(aes(y = 15*avg_rate)) + 
#   scale_y_continuous(sec.axis = sec_axis(~./15))


#Data Visualization for entries from 1990s onward 
recent_movies <- Without_outlier %>% 
  filter(year >= 1990)

data_vis <- recent_movies%>% 
  ggplot(aes(x = year)) + 
  geom_bar(aes(y = count), stat = "Identity") + 
  geom_line(aes(y = 15*avg_rate)) + 
  scale_y_continuous(sec.axis = sec_axis(~./15, name = "Exchange Rate USD/CAD")) + 
  labs(title="Exchange Rate and its Impact on Toronto's Film Industry", x ="year", y = "Count") + 
  theme_economist() + 
  scale_colour_economist() +
  theme(plot.title = element_text(size=12, face="bold.italic"),
        axis.title.x = element_text(size=10, face="bold"),
        axis.title.y = element_text(size=10)) 



```


 
```{r, warning= FALSE, echo= FALSE}
# Add years of recesssion
library(scales)
recession_added <- recent_movies %>%
  mutate(recessions = case_when(
    year >= 1990 & year<= 1991 ~ "recession", 
    year == 2001 ~ "recession", 
    year >= 2007 & year <= 2009 ~ "recession", 
    TRUE ~ "no recession")) %>% 
  mutate_if(is.numeric, replace_na, 0) 

recession_added %>% 
  ggplot(aes(x = year)) + 
  geom_bar(aes(y = count, fill = recessions), stat = "Identity") +  
  scale_fill_manual(name = "Economic Impact", labels = c("Recession", "No Recession"), values = c("recession" = "#cb4b16", "no recession" = "#268bd2")) +
  geom_line(aes(y = 15*avg_rate), colour = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./15, name = "Exchange Rate USD/CAD")) +
  labs(title="Does Exchange Rate and Recessions Affect the Number of Films Made in Toronto?", x ="year", y = "Count") + 
  theme_economist() + 
  scale_colour_economist() +
  theme(plot.title = element_text(size=11, face="bold.italic"),
        axis.title.x = element_text(size=10, face="bold"),
        axis.title.y = element_text(size=10))

#Solarized color palette (discrete)
#e3120b
#4a4a4a

#Economist color_palette 
#cb4b16
#268bd2
```

## Data Visualization Writeup 
The purpose of my visualization graph is to understand whether the ratio of USD/CAD exchange rate have any effect to the number of film productions in Toronto. To better assist the viewers to understand the general economic condition in each year , I also highlighted the years that United States experienced recessions so the viewers can make better comparisons. According to our graph, we see that in the years where exchange rates are low (ex. 1998- 2002), the number of movies made in Toronto is the highest. On the other hand, we don't see a direct effect from the economic conditions from the visualization. This may due to the lag effect from the recessions however, since we haven't done much statistical analysis here, more study is needed before drawing any conclusions. 

My data visualization is a combination of both a line chart and a bar chart. The line chart represents the exchange rate between US dollar(USD) and Canadian dollar(CAD). The bar chart is the number of movies filmed in Toronto, separated by years ranging from 1990-2019. The bar chart is separated by 2 types of economic conditions: years in recession (orange) and years not in recession (blue), and this would be a categorical attribute. The x-axis in the graph is measured by year, an ordinal, sequential data attribute. The y-axis is measured by 2 scales due to the combination of line and bar chart, and they are "the number of movie productions made in Toronto" and "USD/CAD exchange rate". Both attributes are quantitative and sequential attribute types. 

In the process of making this data visualization, I found the dataset also contains data points dated back in the 1930s. If included those points in the graph, it would stretch the x-axis out very long and create unnecessary blank spaces. Therefore, I focus only on the data points between 1990- 2019. Moreover, I imported extra data from https://ca.investing.com/currencies/cad-usd-historical-data for the historical exchange rate between USD to CAD. In the plotting process, I scaled the ratio up by 15 times so the line graph seems more obvious comparing to the number count of the movies. 

The hues I chose to present the economic conditions for each year are blue and orange from the Discrete Solarize Colour Palette. The two hues are very distinctive from each other, since economic conditions are categorical variables. This way, viewers can easily spot the recession years that I highlighted on the graph. 






